\chapter*{Preface}
When I graduated high school with a degree in Latin and English, I could not have imagined that six years later, I would be working in one of today's most rapidly advancing technological fields. And yet, this has become a reality. During my master thesis, I worked together with some of the brightest engineers, biologists and bioinformaticians in the field, on the development of technology which I believe will truly revolutionise biology and improve the lives of many. I am indebted to many people for where I am today, but I am especially grateful to my parents and my brother, who have continuously supported me throughout my career. Moeke, Vake, Freek - I may not always say it out loud, but you all share a special place in my heart. Without you, I would have been nowhere.\pms

Several people were involved in bringing this work to a successful ending. First, I want to thank Valerie and Samira. The two of you helped me in some of the most important experiments performed in this project, even when your time was limited. Second, Gert, Kris and Katina. You provided excellent support in processing our sequencing data, and reassured me that I had not (completely) screwed up the experiments. Stein, Suresh, in working together with you closely for the past year, I have started to look up to you not just as great scientists, but also as great people. Thank you for sharing your vast knowledge and experience with me, and for devoting a large part of your time and patience to this work. You inspire me.\pms

Finally, Jasper. When you were proofreading my thesis right before submission, I showed you a mock version of this preface where you had not received nearly as much praise as you deserve. Joke's on you, I dedicated a whole paragraph to you in the final version. You supported me from day one until the end - when things were going well, and when they were not. We spent many long hours in the lab and in the office, and you answered my panic-driven questions at 2 am in the morning, the weekend before the deadline. You are one of the most extraordinarily kind, intelligent and patient people I have ever met, and I consider you a mentor and a dear friend. Thank you for everything.\pms

\vspace{36pt}
This thesis is dedicated to my biological parents. I do not know where you are, or what you look like, but I love you. Your loss has not been in vain.\pms
